---
title: "analysis_Roman_data"
author: "ls760"
date: "19/04/2021"
output: 
  html_document:
     code_folding: "hide"
     anchor_sections: TRUE
     highlight: "pygments"
     template: "default"
     toc: true
     df_print: paged
     toc_depth: 2
---

```{r, echo=F, message=FALSE, warning=F}

# Set the environment
library(skimr)
library(ggplot2)
library(gghighlight)
library(tidyverse)
library(reshape2)
library(plyr)
library("ggpubr")
library(ggExtra)
```


## Data

```{r read in data, tidy=TRUE,  warning=FALSE, message=FALSE}
dfr = data.table::fread("/Volumes/GoogleDrive/My Drive/Phd/VarioPath/protein_structure/Roman's_VarioPath_scores_20210312.tsv", header = T)
## preview dfr
# str(dfr)
# remove V8 that is empty
dfr = dfr[,-8]
## inspect the DF
skim(dfr)
# Transform variables to numeric
# convert class to numeric where possible
dfr = as.data.frame(dfr)
for(i in 1:length(colnames(dfr))){
  if ( sum(is.na(as.numeric(as.character(dfr[,i])))) == dim(dfr)[1] ) {
    message(colnames(dfr)[i], ' not converted')
  } else {
    dfr[,i] = as.numeric(as.character(dfr[,i]))
  }
}


```

## MDT Variants

Are MDT variants similar to the rest of the cohort?

```{r MDT_All_variants, warning=FALSE, message=FALSE}
ggplot(data = dfr, mapping = aes(Score, fill = MDT_variant))+
  geom_histogram(bandwidth = 1, bins = 40, na.rm = TRUE) 

ggplot(data = dfr, mapping = aes(Score, fill = MDT_variant))+
  geom_density(na.rm = TRUE) 

ggplot(data = dfr, mapping = aes(Category, fill = MDT_variant))+
  geom_histogram(na.rm = TRUE) 

```

## How does romans score correlates with effect size

This analysis is difficult because Romans' score doesn't take into account the phenotype, but is intrinsi to the protein. However, effect size strongly rely on the phenotype taken into account.

I will take the max effect size reported for that variant. assuing that Roman's score predict the worst outcoma as the highest score. However, there is a chance that my varaints haven't been tested against the right phenotype.

Effect size can be positive or negative, so I'll use *effect_size^2^*

I'll look one gene at the time on superimpose the different genes


```{r correlation_roman_effect_size, warning=FALSE, message=FALSE}
# import effect sizes
dir_es = "/Users/luca/Library/Group Containers/G69SCX94XU.duck/Library/Application Support/duck/Volumes/logincpu.hpc.cam.ac.uk – SFTP/UKBb/variopath/burden_test_raremetal/output_pre_MDT/phenotypes"

list_files_scores = list.files(path = dir_es, pattern = "*.dominant.singlevar.score.txt.gz$", full.names = TRUE, recursive = TRUE)

# Read and create reference table with all effect sizes
data_all =  
  lapply(list_files_scores, data.table::fread, skip = "#CHROM", header = TRUE) %>%
  bind_rows(.id = "file")

skim(data_all)

sign_pvalue = 0.05

df_eff = data.table::dcast(data_all, 
                  `#CHROM` + POS + REF + ALT ~ file, 
                  value.var = c('ALT_EFFSIZE', "PVALUE"), function(x){max(x,na.rm = T)},
                  subset = .(PVALUE < sign_pvalue & is.na(PVALUE) == FALSE))

# return only the max squared effect size per variant
df_eff = as.data.frame(df_eff)

df_eff[sapply(df_eff, is.infinite)] <- NA

df_eff$sqr_eff = sqrt(apply(df_eff[,grep(colnames(df_eff) ,pattern = "ALT_EFFSIZE*")], 
      1, 
      function(x){
        max(x^2, na.rm = T)
      }))

df_eff$id = paste(paste0( 'chr', df_eff$`#CHROM`), 
                  df_eff$POS, df_eff$REF, df_eff$ALT, sep="_")

googledrive::drive_download(file = googledrive::as_id('https://docs.google.com/spreadsheets/d/1OBQmTN5BVhRgdXh-jM3OxExd6WxwJP_MYflJTg3dNUM/edit#gid=1156788760'))
gene_list_moi = readxl::read_xlsx('VarioPath_disease_gene_list_20210514v1.0.xlsx')
file.remove('VarioPath_disease_gene_list_20210514v1.0.xlsx')


gel = unique(gene_list_moi$`MISC_MOIs (GEL)`)
omes = unique(gene_list_moi$`MISC_MOIs (OMIM via ENSG)`)
nbr = unique(gene_list_moi$`MISC_MOIs (NBR)`)

revalue_vector = c("AD","AR")
names(revalue_vector) = c("Autosomal dominant", "Autosomal recessive")

gene_list_moi$`MISC_MOIs (OMIM via ENSG)` = revalue(gene_list_moi$`MISC_MOIs (OMIM via ENSG)`, replace = revalue_vector, warn_missing = T)

gene_list_moi$moi = apply(gene_list_moi[,c(23,25)], 
      1,
      function(x)
            stringi::stri_remove_empty(
              str_trim(paste(unique(str_trim(str_split(x, "\\,|\\;|\\|",simplify = T))), collapse = " "), "both")
                     )
      )

pli= data.table::fread("~/Desktop/VarioPath/disease_gene_list_20200915_pLI_info.csv")

gene_pli_moi = merge(pli, gene_list_moi, by.x = "gene", by.y = "Approved symbol (HGNC)") %>% select(c("gene", "moi", "pLI"))

df_cor = merge(df_eff, dfr, by.x = 'id', by.y = 'Uploaded_variation', all.x = T) %>% 
  select(c(Score,sqr_eff, SYMBOL)) 

df_cor = merge(df_cor, gene_pli_moi, by.x = "SYMBOL", by.y = "gene")

df_cor=xlsx::read.xlsx("data_to_check.xlsx", sheetIndex = 1)


p = ggscatter(df_cor, x = "Score", y = "sqr_eff",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Effect Size", ylab = "SVM protein score")
ggMarginal(p, type = 'densigram')

p = ggscatter(df_cor, x = "Score", y = "sqr_eff", color = "pLI", fill = "pLI",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Effect Size", ylab = "SVM protein score")
ggMarginal(p, type = 'densigram')

p = ggscatter(df_cor, x = "Score", y = "sqr_eff", color = "moi", fill = "moi",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Effect Size", ylab = "SVM protein score")
ggMarginal(p, type = 'densigram')


cor.test(df_cor$Score, df_cor$sqr_eff, method = "pearson",use = 'pairwise.complete.obs')

# Color code mode of inheritance and pLI




```




```{r correlation_roman_effect_size_after_MDT, warning=FALSE, message=FALSE}
# import effect sizes
dir_es = "/Users/luca/Library/Group Containers/G69SCX94XU.duck/Library/Application Support/duck/Volumes/logincpu.hpc.cam.ac.uk – SFTP/home/ls760/UKBb/variopath/burden_test_raremetal/output_after_MDT_clean/phenotypes/"

list_files_scores = list.files(path = dir_es, pattern = "*.dominant.singlevar.score.txt.gz$", full.names = TRUE, recursive = TRUE)

# Read and create reference table with all effect sizes
data_all =  
  lapply(list_files_scores, data.table::fread, skip = "#CHROM", header = TRUE) %>%
  bind_rows(.id = "file")

skim(data_all)

sign_pvalue = 0.05

df_eff = data.table::dcast(data_all, 
                  `#CHROM` + POS + REF + ALT ~ file, 
                  value.var = c('ALT_EFFSIZE', "PVALUE"), function(x){max(x,na.rm = T)},
                  subset = .(PVALUE < sign_pvalue & is.na(PVALUE) == FALSE))

# return only the max squared effect size per variant
df_eff = as.data.frame(df_eff)

df_eff[sapply(df_eff, is.infinite)] <- NA

df_eff$sqr_eff = sqrt(apply(df_eff[,grep(colnames(df_eff) ,pattern = "ALT_EFFSIZE*")], 
      1, 
      function(x){
        max(x^2, na.rm = T)
      }))

df_eff$id = paste(paste0( 'chr', df_eff$`#CHROM`), 
                  df_eff$POS, df_eff$REF, df_eff$ALT, sep="_")

df_cor = merge(df_eff, dfr, by.x = 'id', by.y = 'Uploaded_variation', all.x = T) %>% 
  select(c(Score,sqr_eff)) 

stats::cor(df_cor$Score, df_cor$sqr_eff, method = "pearson",use = 'pairwise.complete.obs')

# Color code mode of inheritance and pLI

p = ggplot(df_cor, aes(Score, sqr_eff)) + 
  geom_point() 

ggMarginal(p, type = 'densigram')
```
