---
title: "check_gender_distribution_xlinked"
author: "ls760"
date: "11/02/2021"
output:
  pdf_document: 
    fig_crop: no
    fig_height: 13
    fig_width: 12
  html_document:
    fig_height: 10
    fig_width: 14
    highlight: tango
    theme: readable
---

## Gender distribution of X-linked diseases

I selected the genes that are cause of X-linked diseases (plus NBEAL2 as control):

| Gene | disease | MOI |
| ----- | ------- | ---- |
| F8 | FVIII deficiency | XLR |
| F9 | FIX deficiency | XLR
| FLNA | thrombocytopenia with giant platelets |   |
| NBEAL2 | GPS  |  Not x-linked. used as control    
| PIGA | Paroxysmal nocturnal hemoglobinuria  | Somatic | 
| WAS | Wiskott-Aldrich syndrome | XLR |

```{r, echo=T, warning=F, message=F}
require(reshape2)
require(tidyverse)
load('variopath_MDT.RData')
# create variant IDs
df_clean$ID = paste(df_clean$CHROM, df_clean$POS, df_clean$REF, df_clean$ALT, sep = '_') 
# reshape the table to have long format
df_gender = unique(df_gender)
df_gender = melt(df_clean[ (df_clean$CHROM == 'X' | df_clean$GENE == 'NBEAL2'), 
                           c('ID','GENE','male','female') ])
```

```{r, echo=T, warning=F, message=F}
# Check where are the majority of x-linked variants in male
ggplot(df_gender, aes(x = ID, 
                      y = value, 
                      fill = variable) ) +
  geom_bar(stat = 'identity', position = 'dodge2') + 
  facet_wrap('GENE', scales = 'free') + 
  theme(axis.text.x = element_text(angle = 90))
```

## Stats gene-level
Look at how many people are hemizigous/male in x-linked genes

```{r, echo=T, warning=F, message=F}
gene_list  = df_clean[( df_clean$CHROM == 'X' | df_clean$GENE == 'NBEAL2'),
                      c('GENE','male','female') ] %>% 
  group_by(GENE) %>% 
  summarise('n_male' = sum(male),'n_female' = sum(female))
knitr::kable(gene_list, format = "pipe")
```

### Fisher exact test

```{r, echo=T, warning=F, message=F}
gene_list$fisher.pvalue = "1"
for (gene in 1:dim(gene_list)[1]) {
  total_obs_gender = (gene_list[gene,2] + gene_list[gene,3]) / 2 
  # assuming 50:50 distribution between the genders
  fisher.matrix = matrix(c(as.integer(gene_list[gene,2]),
                           as.numeric(gene_list[gene,3]), 
                           as.numeric(total_obs_gender), 
                           as.numeric(total_obs_gender) ), 
                         nrow = 2, 
                         ncol = 2,
                         byrow = T, 
                         dimnames =
                           list(c("observed", "expected"),
                                c("male", "female")))
  gene_list$fisher.pvalue[gene] = fisher.test(fisher.matrix)$p.value
}
knitr::kable(gene_list, format = 'pipe')
```

### Example of a fisher matrix

```{r, echo=F, warning=F, message=F}

#Examlpe matrix
knitr::kable(gene_list[gene,1],format = 'simple')
knitr::kable(fisher.matrix,format = 'simple')
```

# --- End of the script ---



```{r, echo=F, warning=F, message=F}
sessionInfo()
```









